---
title: "Introduction to Programming"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)

```


```{r loading-data, context = "data"}
ice_cream <- read.csv("./www/ice_cream_sales.csv")

```

## Programming Basics

All programming languages (e.g., R, Python, HTML, CSS, etc.) share similar components. We will start by learning about the basics of programming, before digging into our data.

### Comments

Comments are an integral part of the coding process. Comments are notes about what the code does, how it works, or to create sections (like using headers in a word doc). Comments help remember what your code does and helps other people understand your code when you share it.

-   In R, comments begin with a \#
    -   The pound, hash, or hashtag symbol tells R that the line that starts with a \# should not be run like the rest of the code.e
-   In this workshop, \## indicates comments.
    -   These lines should not be run and the \## should not be deleted.
-   A single \# will be for lines of code that are commented out, but will run if you delete the #.

```{r comment-example, exercise = TRUE}

## this line will cause R to error if you delete the ##

# print("If you delete the # on this line, it will run!")


```

### Variables

The second part of all coding languages are **variables.** Variables are like boxes. Inside a variable we can store different values. For example, if we wanted to find the sum of 5 and 6, in the R console we can type `5+6` and hit enter. R will return `11`.

R can also do more complicated math like adding and multiplying a bunch of numbers.


```{r math-example, exercise = TRUE}

5+6+8+65654*7846561+94651-15

```


These numbers and the result (515158210549) are not stored anywhere. We can't access these values again if we wanted them. We would have to run the code again if we wanted to get the same answer. That's not super efficient.

Instead, we're better off using **variables**.
- Variables are like containers or boxes. They can be empty or they can store something
  - In R, when we want to create a variable, we need to use the assignment operator: `<-`. 
  - **NOTE:** Throughout these online tutorials, you'll see the assignment operator using: `<<-`
    - with two left-angle brackets `<<`. This allows variables to be accessible across code chunks.
  - on the left side of the operator is the variable name, on the right side of the operator is the value we want to store
    - `num_five <- 5`
    - This stores the integer `5` in the variable `num_five`.
- When we store values into variables we can use the variables instead of the values to do different tasks.

```{r variable-example, exercise=TRUE}
## store the integer 5 into the variable num_five
num_five <- 5

## add num_five to 6
num_five + 6
```

Variables can also hold different **data types**

### Data Types

There are many different data **types**. To save time and space, I'll only over the data types we'll use the most.

#### characters or strings
- characters or strings are words.
- In code they're surrounded by quotation marks.
- `"Hello!"` is a string where `Hello!` is not.
- It's important to know what type of data you have because if you have numbers stored as characters, you can't do math using strings. 
  - so `"5"` is a string while `5` is an integer.

```{r setupA}
char_five <- "5"
int_five <- 5
```

  
```{r characters, echo=TRUE, error=TRUE, message=TRUE, exercise=TRUE}

## store five as an integer
int_five <- 5

## the following line will work
int_five + 6

## store five as a character
char_five <- "5"

## the following line will give an error
# char_five + 6
```

In the code above, delete the \# before line 11 and run the code again. You will get an error that says `non-numeric argument to binary operator`. That's a fancy way of saying R does not know how to add the character `"5"` to the integer `6`. It would be like adding `a` to `6`.

#### numbers

So how do we store numbers? There are two main ways to store numeric values in R. 
- **doubles** are numbers that have decimal points.
  - `10.5`, `1241.22314`, and `56.0` are all doubles.
- **integers** are whole numbers
  - `1`, `12314151`, and `54` are all integers.

#### lists

Lists are vectors that contain other values. They can store strings, integers, doubles, and other variables. 

#### data frames

A data frame is R's version of a spreadsheet. It has columns and rows and can store a variety of data types.

#### Finding an Object's Class

R has a built in function to check what class (or data type) an object is. You can use the `class()` function.


```{r class-example, exercise=TRUE, exercise.setup="setupA"}

## use the class function to check data types
class(char_five)
class(int_five)

```

#### Summary

For the most part, you won't need to worry about what type of data you'll be using. It only becomes an issue when you need to combine data together. 

As demonstrated above, you can't combine strings and integers. We'll cover how to fix this as it comes up while working with data.

### Functions

Functions are blocks of codes that only run when called. We'll be using built-in functions to add and explore data. For example, let's look at the `head()` function. `head()` will return the first five rows of a data set. 

R has several built-in data sets. We're going to look at the first five rows of the `cars` data set.

```{r functions-intro, exercise=TRUE}
head(cars)
```


Functions usually take **arguments**, which are additional instructions. In the example above, the argument `cars` is the name of the data set we want to get the `head()` of.

## Getting and Setting the Working Directory  

Before we can load data, we need to know what folder (also known as a directory) the computer is currently in. We do this by using the `getwd()` function. 

```{r get-wd, exercise = TRUE}
getwd()
```

The `getwd()` function returns the path to the current folder we are in. We can change it by using `setwd("Path/to/directory")` and changing the file `path`.

The `path` is part of the code between the quotation marks `"`. It points to where the data is stored on your computer. 

The `path` to the file can be **absolute** or **relative**. 

- An **absolute** path is like giving step by step directions from your house to the store. 
  - For example: `C:/Users/Liz/Documents/Data/` is an absolute path.
  - It tells R that on the `C:/` drive, look in the `Users` folder, then inside the `Bortimus` folder, you'll find the `Documents` folder, and finally the data will be inside the `Data` folder.

More often than not, we'll use **relative** directories.

- A **relative** directory is like telling people that the store is two blocks from your house, or near another landmark.
  - For example: `./Data/` is a relative directory that tells R to look in the current directory (folder) for the `Data` folder.
  - `../Data/` is another relative directory that tells R in that we should look for the Data folder in the parent directory of where we are on our computer. 
```{r setting-wd, eval=FALSE, echo=TRUE}
setwd("C:/Users/Liz/Documents/Data")
```

If you're running this on a web browser, you won't need to change anything. The data is already loaded for you.

If you're using this code on your computer, just change the part in the quotation marks to where the data is stored on your computer.

## Loading and Examining Data

In this section, we will learn how to load and work with data. R can load many different file types. We will primarily use `.csv` files, which are comma separated values. These are normal spreadsheets. 

We're going to start with a very basic csv with 500 rows. 

The first step is to load the data. We do this by creating a **variable** to store the data once it's loaded. We also use a default function in R `read.csv()`.

The `read.csv()` function reads data stored in common separated value (`.csv`) files.

```{r load-ice-cream, echo=TRUE}
ice_cream <- read.csv("./www/ice_cream_sales.csv")
```

You will need to change the `path` or the part between the quotation marks `"` to the location of the data. The data is available on the Google Drive under `Data/practice`.

If you are using the web version of this tutorial, you do not need to change anything. The data is already loaded into the environment.

#### inspecting the data

We'll start by getting the names of the columns in our data set. We do this by using the `colnames()` function. Inside the `colnames` function, we need to tell R what data set to get the column names of, so we'll pass `ice_cream` as the argument.


In the space below, find the column names for the `ice_cream` data 
```{r column-names, exercise=TRUE}

```

Our data has two columns:

- `Temperature`: This column lists the temperature in degrees celsius.
- `Revenue`: This column is the amount of ice cream sold, in dollars.

#### summarizing the data

We can also get the summary of the data. The `summary()` function in R will give the minimum, maximum, median, mean, and value of first quartile (25%) and third quartile (75%)

In the space below, use the `summary()` function to get an overview of our data by passing `ice_cream` as the argument.

```{r summary-ice-cream, exercise=TRUE}

```

#### head & tail

Another good way to look at our data is to view the head or tail. `head()` returns the top six rows and `tail` the bottom six. We can also include `n = #` to tell R to get more or fewer rows.

In the space below, pass `ice_cream` as the argument for `head()` and `tail()`. 

If you want to change how many rows are returned, use `head(<data-set-name>, n = 10)` 

```{r head-tail, exercise=TRUE}

```












































