---
title: "Introduction to Programming and Statistics"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)


```

load data

```{r loading-data, context = "data"}
elec_demo <- read.csv("./www/election_demographics.csv")

```

## Programming Basics

All programming languages (e.g., R, Python, HTML, CSS, etc.) share similar components. We will start by learning about the basics of programming, before digging into our data.

### Comments

Comments are an integral part of the coding process. Comments are notes about what the code does, how it works, or to create sections (like using headers in a word doc). Comments help remember what your code does and helps other people understand your code when you share it.

-   In R, comments begin with a \#
    -   The pound, hash, or hashtag symbol tells R that the line that starts with a \# should not be run like the rest of the code.e
-   In this workshop, \## indicates comments.
    -   These lines should not be run and the \## should not be deleted.
-   A single \# will be for lines of code that are commented out, but will run if you delete the #.

```{r comment-example, exercise = TRUE}

## this line will cause R to error if you delete the ##

# print("If you delete the # on this line, it will run!")


```

### Variables

The second part of all coding languages are **variables.** Variables are like boxes. Inside a variable we can store different values. For example, if we wanted to find the sum of 5 and 6, in the R console we can type `5+6` and hit enter. R will return `11`.

R can also do more complicated math like adding and multiplying a bunch of numbers.


```{r math-example, exercise = TRUE}

5+6+8+65654*7846561+94651-15

```


These numbers and the result (515158210549) are not stored anywhere. We can't access these values again if we wanted them. We would have to run the code again if we wanted to get the same answer. That's not super efficient.

Instead, we're better off using **variables**.
- Variables are like containers or boxes. They can be empty or they can store something
  - In R, when we want to create a variable, we need to use the assignment operator: `<-`. 
  - on the left side of the operator is the variable name, on the right side of the operator is the value we want to store
    - `num_five <- 5`
    - This stores the integer `5` in the variable `num_five`.
- When we store values into variables we can use the variables instead of the values to do different tasks.

```{r variable-example, exercise=TRUE}
## store the integer 5 into the variable num_five
num_five <- 5

## add num_five to 6
num_five + 6
```

Variables can also hold different **data types**

### Data Types

There are many different data **types**. To save time and space, I'll only over the data types we'll use the most.

#### characters or strings
- characters or strings are words.
- In code they're surrounded by quotation marks.
- `"Hello!"` is a string where `Hello!` is not.
- It's important to know what type of data you have because if you have numbers stored as characters, you can't do math using strings. 
  - so `"5"` is a string while `5` is an integer.
  
```{r characters, echo=TRUE, error=TRUE, message=TRUE, exercise=TRUE}

## store five as an integer
int_five <- 5

## the following line will work
int_five + 6

## store five as a character
char_five <- "5"

## the following line will give an error
# char_five + 6
```

In the code above, delete the \# before line 11 and run the code again. You will get an error that says `non-numeric argument to binary operator`. That's a fancy way of saying R does not know how to add the character `"5"` to the integer `6`. It would be like adding `a` to `6`.

#### numbers

So how do we store numbers? There are two main ways to store numeric values in R. 
- **doubles** are numbers that have decimal points.
  - `10.5`, `1241.22314`, and `56.0` are all doubles.
- **integers** are whole numbers
  - `1`, `12314151`, and `54` are all integers.

#### lists

Lists are vectors that contain other values. They can store strings, integers, doubles, and other variables. 

#### data frames

A data frame is R's version of a spreadsheet. It has columns and rows and can store a variety of data types.

#### Finding an Object's Class

R has a built in function to check what class (or data type) an object is. You can use the `class()` function.

```{r class-example, exercise=TRUE, dependson=-1}

## use the class function to check data types
class(char_five)
class(int_five)

```

#### Summary

For the most part, you won't need to worry about what type of data you'll be using. It only becomes an issue when you need to combine data together. 

As demonstrated above, you can't combine strings and integers. We'll cover how to fix this as it comes up while working with data.

















