---
title: "Introduction to Statistics pt. II: Controls & Interactions"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```

```{r loading-data, context = "data"}
elec_demo <- read.csv("./www/election_demographics.csv") %>% 
  mutate(DVP = (DEM/total_votes)*100,
         RVP = (REP/total_votes)*100,
         log_hhic = log(med_hhic))
```

## Linear Regression & Controls

In [Introduction to Statistics](https://liz-muehlmann.shinyapps.io/intro_stats/) we began looking at control variables.

Controls are variables that are held constant throughout the experiment. Control variables may influence the outcome, but they aren't what we're interested in primarily. 

We are going to work with our data set `elec_demo`. 

Let's load the data (*reminder*: if you're using the web version, the data is already loaded):

```{r load-data, echo=TRUE}
elec_demo <- read.csv("./www/election_demographics.csv") %>% 
  mutate(DVP = (DEM/total_votes)*100,
         RVP = (REP/total_votes)*100,
         log_hhic = log(med_hhic))
```


A linear regression in R uses the following form: `lm(Y ~ X, data = "data_set_to_use")`. In English, this translates to `linear model(Response ~ Predictor, data = "data_set_to_use)`. When we want to add in control variables, we add them using the `+` operator and then the variable we want to use as a control. The form then becomes `lm(Y ~ X + Z, data = "data_set_to_use")` where `Z` is our control variable. 

We're going to use the `elec_demo` data to add in controls to our linear regression model. 

Let's predict Democratic Voting Percentage (`DVP`) using `median gross rent`.

In the space below, use the variable `med_grossRent` to predict Democratic Voting Percentage, `DVP`.

```{r lm-dvp, exercise=TRUE}

```

```{r lm-dvp-hint}

## run the linear regression
## DV: Democratic Voting Percentage
## IV: Median Gross Rent

lm_dvp <- lm(DVP ~ med_grossRent, data = elec_demo)

## get the summary output for the regression

summary(lm_dvp)
```


In this model, we can see that as median gross rent increases by $100, democratic voting percentage increases by 0.04.


## Adding in a Control

```{r lm-dvp-control, exercise=TRUE}

```

```{r lm-control-hint}

## run the linear regression
## DV: Democratic Voting Percentage
## IV: Median Gross Rent
## Control: Percent of the population 

lm_dvp <- lm(DVP ~ med_grossRent, data = elec_demo)

## get the summary output for the regression

summary(lm_dvp)
```



The first step in interpreting the model with multiple predictor variables (here, control, Hawthorne, and Neighbors) is to look at the *f-statistic* at the bottom of the model. Our *f-statistic* is < 0.05, which means that at least one of our predictors is statistically significant. 

For each of our predictors, we can interpret the coefficient as the average effect on whether a person voted in the 2006 primary election while holding all other predictors fixed. 
























